---
title: "HW2_McGill_Max"
author: "Max McGill"
date: 2017-09-11
output:
  pdf_document: default
subtitle: Due Monday September 11, 10 am
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

\section{Problem 4}
In the classroom, version control can assist in:
\begin{itemize}
\item updating and reverting code as necessary for assignments and projects
\item maintaining and securing multiple versions of code for comparison, reversion, and conversion
\item facilitating collaborative efforts for projects that require such methods
\end{itemize}
These functionalities benefit a classroom environment that makes heavy use of programming methods by providing the tools necessary to alter code, either alone or collaboratively, in such a way that facilitates an ease of maneuverability between versions of the code developed or used.
In turn, this provides access to a host of options for the modification and sharing of code that simplify the classroom experience of tasks.

\section{Problem 5}
\subsection{Part A}
```{r,echo=FALSE}
urla<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensoryraw<-read.table(urla, header=F, skip=1, fill=T, stringsAsFactors = F)
sensorytidy<-sensoryraw[-1,]
#The data is recorded in triplicate, resulting in missing sets of item values when read.
#This causes a shift of the values in rows with an integer and those without.
#Create two dataframes by filtering rows with integers in V1.
#Rename columns in the wone with integer values appropriatey.
sensorytidy_a<-filter(.data = sensorytidy,V1 %in% 1:10) %>%
                rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
#In the one without, create a column to represent missing item values.
sensorytidy_b<-filter(.data = sensorytidy,!(V1 %in% 1:10)) %>%
                mutate(Item=rep(as.character(1:10),each=2)) %>%
                mutate(V1=as.numeric(V1)) %>%
                select(c(Item,V1:V5))
#bind the dataframes and name their columns as relevant. 
sensorytidy<-bind_rows(sensorytidy_a,sensorytidy_b)
colnames(sensorytidy)<-c("Item",paste("Person",1:5,sep="_"))
#Gather and mutate the person variable to create a numerically valued column.
#Arrange the data by item.
sensorytidy<-sensorytidy %>%  
    gather(Person,value,Person_1:Person_5) %>%  
    mutate(Person = gsub("Person_","",Person)) %>%
    arrange(Item)
```
```{r echo=F, eval=T, include=T}
knitr::kable(summary(sensorytidy),caption="syntax may be slightly off")
```

\subsection{Part B}
```{r,echo=FALSE}
urlb<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
goldraw<-read.table(urlb,header=T,fill=T,stringsAsFactors=F)
#The data are split among multiple columns that must be combined into coherence.
#Create new dataframes for each variable by selecting their resective columns.
#Then gather their columns into new data frames to create a singular column for each.
yeardf<-select(goldraw,"Year","Jump","Long.1","Year.2")
yearvalue<-gather(yeardf,Year:Year.2)
longjumpdf<-select(goldraw,"Long","Year.1","Jump.1","Long.2")
longjumpvalue<-gather(longjumpdf,Long:Long.2)
#Combine these combined columns to create a new dataframe.
goldna<-data.frame(yearvalue$value,longjumpvalue$value)
#Remove the nonexistent values to create the tidy data frame.
goldtidy<-drop_na(goldna)
#Set the variable names to appropriate values.
colnames(goldtidy)<-c("Year","Long Jump")
#Mutate the year variable to display the years represented.
goldtidy<-mutate(goldtidy, Year=Year+1900)
summary(goldtidy)
```
\subsection{Part C}
```{r,echo=FALSE}
urlc<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brainbodyraw<-read.table(urlc, header=F, skip=1, fill=T, stringsAsFactors = F)
#This data contains two variables spread across multiple columns.
#Rename the columns using repetiton to properly designate the variables.
colnames(brainbodyraw)<-rep(c("Brain","Body"),3)
#Create a new dataframe by binding the individual pairs of columns.
brainbodytidy<-rbind(brainbodyraw[,1:2],brainbodyraw[,3:4],
                         brainbodyraw[,5:6])
#Remove nonexistent values.
brainbodytidy<-brainbodytidy %>%  
    filter(!(is.na(Brain))) 
summary(brainbodytidy)
```
\subsection{Part D}
```{r,echo=FALSE}
urld<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomatoraw<-read.table(urld, header=F, skip=2, fill=T, stringsAsFactors = F, comment.char =
                        "")
#The data is listed in triplicate by commas, contains an extraneous comma in one variable,
#and contains a character hiding the first row of data.
tomatotidy<-tomatoraw %>%  
#Separate each collection of observations into appropriate columns. 
    separate(V2,into=paste("C10000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
    separate(V3,into=paste("C20000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
    separate(V4,into=paste("C30000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
#Remove the extraneous comma.
    mutate(C10000_3=gsub(",","",C10000_3)) %>%
#Gather columns and rename the variables as appropriate.
    gather(Clone,value,C10000_1:C30000_3) %>%
    mutate(Variety=V1, Clone=gsub("C","",Clone)) %>%
    mutate(Variety=gsub("\\\\#"," ",Variety)) %>%
#Separate the clone column into variables clone and replicate.
    separate(Clone,into = c("Clone","Replicate")) %>%
#Exclude the V1 column and arrange by variety.
    select(-V1,Variety,Clone,value) %>%
    arrange(Variety)
#All variables are still listed as characters.
summary(tomatotidy)
```
\section{Problem 6}
```{r,echo=FALSE, eval=FALSE}
#Path to data
.datapath <- file.path(path.package('swirl'), 'Courses',
                      'R_Programming_E', 'Looking_at_Data',
                      'plant-data.txt')
# Read in data
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")
# Remove annoying columns
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% .cols2rm)]

# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period',
                   'Foliage_Color', 'pH_Min', 'pH_Max',
                   'Precip_Min', 'Precip_Max',
                   'Shade_Tolerance', 'Temp_Min_F')
plants<-mutate(plants,pH_Range=pH_Max-pH_Min)
plants<-select(plants,c('Foliage_Color','pH_Range'))
plants<-drop_na(plants)
plants<-arrange(plants, Foliage_Color)
#Can't figure out howw to convert it.
plants.anova<-lm(pH_Range ~ Foliage_Color, data=plants)
```
\section{Problem 7}
```{r,echo=FALSE}

```
\section{Appendix 1: R Code}
\subsection{Problem 5}
```{r,eval=FALSE}
#Part A
urla<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensoryraw<-read.table(urla, header=F, skip=1, fill=T, stringsAsFactors = F)
sensorytidy<-sensoryraw[-1,]
#The data is recorded in triplicate, resulting in missing sets of item values when read.
#This causes a shift of the values in rows with an integer and those without.
#Create two dataframes by filtering rows with integers in V1.
#Rename columns in the wone with integer values appropriatey.
sensorytidy_a<-filter(.data = sensorytidy,V1 %in% 1:10) %>%
                rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
#In the one without, create a column to represent missing item values.
sensorytidy_b<-filter(.data = sensorytidy,!(V1 %in% 1:10)) %>%
                mutate(Item=rep(as.character(1:10),each=2)) %>%
                mutate(V1=as.numeric(V1)) %>%
                select(c(Item,V1:V5))
#bind the dataframes and name their columns as relevant. 
sensorytidy<-bind_rows(sensorytidy_a,sensorytidy_b)
colnames(sensorytidy)<-c("Item",paste("Person",1:5,sep="_"))
#Gather and mutate the person variable to create a numerically valued column.
#Arrange the data by item.
sensorytidy<-sensorytidy %>%  
    gather(Person,value,Person_1:Person_5) %>%  
    mutate(Person = gsub("Person_","",Person)) %>%
    arrange(Item)
summary(sensorytidy)
```

```{r,eval=F}
#Part B
urlb<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
goldraw<-read.table(urlb,header=T,fill=T,stringsAsFactors=F)
#The data are split among multiple columns that must be combined into coherence.
#Create new dataframes for each variable by selecting their resective columns.
#Then gather their columns into new data frames to create a singular column for each.
yeardf<-select(goldraw,"Year","Jump","Long.1","Year.2")
yearvalue<-gather(yeardf,Year:Year.2)
longjumpdf<-select(goldraw,"Long","Year.1","Jump.1","Long.2")
longjumpvalue<-gather(longjumpdf,Long:Long.2)
#Combine these combined columns to create a new dataframe.
goldna<-data.frame(yearvalue$value,longjumpvalue$value)
#Remove the nonexistent values to create the tidy data frame.
goldtidy<-drop_na(goldna)
#Set the variable names to appropriate values.
colnames(goldtidy)<-c("Year","Long Jump")
summary(goldtidy)
```

```{r,eval=FALSE}
#Part C
urlc<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brainbodyraw<-read.table(urlc, header=F, skip=1, fill=T, stringsAsFactors = F)
#This data contains two variables spread across multiple columns.
#Rename the columns using repetiton to properly designate the variables.
colnames(brainbodyraw)<-rep(c("Brain","Body"),3)
#Create a new dataframe by binding the individual pairs of columns.
brainbodytidy<-rbind(brainbodyraw[,1:2],brainbodyraw[,3:4],
                         brainbodyraw[,5:6])
#Remove nonexistent values.
brainbodytidy<-brainbodytidy %>%  
    filter(!(is.na(Brain))) 
summary(brainbodytidy)
```

```{r,eval=FALSE}
urld<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomatoraw<-read.table(urld, header=F, skip=2, fill=T, stringsAsFactors = F, comment.char =
                        "")
#The data is listed in triplicate by commas, contains an extraneous comma in one variable,
#and contains a character hiding the first row of data.
tomatotidy<-tomatoraw %>%  
#Separate each collection of observations into appropriate columns. 
    separate(V2,into=paste("C10000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
    separate(V3,into=paste("C20000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
    separate(V4,into=paste("C30000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
#Remove the extraneous comma.
    mutate(C10000_3=gsub(",","",C10000_3)) %>%
#Gather columns and rename the variables as appropriate.
    gather(Clone,value,C10000_1:C30000_3) %>%
    mutate(Variety=V1, Clone=gsub("C","",Clone)) %>%
    mutate(Variety=gsub("\\\\#"," ",Variety)) %>%
#Separate the clone column into variables clone and replicate.
    separate(Clone,into = c("Clone","Replicate")) %>%
#Exclude the V1 column and arrange by variety.
    select(-V1,Variety,Clone,value) %>%
    arrange(Variety)
#All variables are still listed as characters.
summary(tomatotidy)
```
\subsection{Problem 6}
\subsection{Problem 7}
