---
title: "STAT 5014 HW4"
author: "Max McGill"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
    rm(list = ls())   
    graphics.off()
    options(scipen = 4, digits = 6)
    options(stringsAsFactors = FALSE)
    knitr::opts_chunk$set(echo = F, eval=T, cache=T, tidy.opts=list(width.cutoff=55),
                tidy=T, include=FALSE, message=F, warning=F)
    library.warn <- library
    library <- function(package, help, pos = 2, lib.loc = NULL, character.only = FALSE,
                        logical.return = FALSE, warn.conflicts = FALSE, quietly = TRUE,
                        verbose = getOption("verbose")) {
       if (!character.only) {
          package <- as.character(substitute(package))
       }
       suppressPackageStartupMessages(library.warn(
          package, help, pos, lib.loc, character.only = TRUE,
          logical.return, warn.conflicts, quietly, verbose))}
```

\section{Problem 3}
According to Roger Peng's \textit{Exploratory Data Analysis with R}, the focus of EDA "is to show the data, summarize the evidence and identify interesting patterns while eliminating ideas that likely won't pan out" (2016, p. 2). 
\section{Problem 4}
Factors within this dataset are limited to the numerical designation by block, while depth and phosphate are continuous, quantitative observations.
The lesson from this exploration is to always loo over your data to ensure there are no lurking dinosaurs, cookie cutter shapes, or other obscuring fabrications.
```{r, echo=FALSE, include=TRUE, fig.width=6, fig.height=4}
#get data
library(xlsx)
prob4_data1 <- read.xlsx("HW4_data.xlsx", sheetIndex = 1)
prob4_data2 <- read.xlsx("HW4_data.xlsx", sheetIndex = 2)
#combine data
prob4data<-rbind(prob4_data1,prob4_data2)

#summarize data
knitr::kable(summary(prob4data), caption="Data Summary")

library(ggplot2)

#from "http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/"
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

#begin plotting the data as a whole
#first a scatter plot
scatter4<-ggplot(prob4data,aes(x=depth,y=phosphate,colour=factor(block))) +
    geom_point() +
    xlab("Depth") +
    ylab("Phosphate") +
    ggtitle("Phosphate by Depth") +
    guides(col=guide_legend(title="Block"))

#to better understand the distributions, a histogram for both variables of interest
dhist4<-ggplot(prob4data, aes(x=depth,fill=factor(block))) +
    geom_histogram(colour="black",binwidth=5) +
    xlab("Depth") +
    ylab("Frequency") +
    ggtitle("Depth for Blocks") +
    guides(fill=guide_legend(title="Block"))
phist4<-ggplot(prob4data, aes(x=phosphate,fill=factor(block))) +
    geom_histogram(colour="black",binwidth=5) +
    xlab("Phosphate") +
    ylab("Frequency") +
    ggtitle("Phosphate for Blocks") +
    guides(fill=guide_legend(title="Block"))

#similar in purpose, a boxplot for both variables of interest
dbox4<-ggplot(prob4data, aes(x=factor(block),y=depth,group=block,fill=factor(block))) +
    geom_boxplot() +
    xlab("Block") +
    ylab("Depth") +
    ggtitle("Depth by Block") +
    guides(fill=guide_legend(title="Block"))
pbox4<-ggplot(prob4data, aes(x=factor(block),y=phosphate,group=block,fill=factor(block))) +
    geom_boxplot() +
    xlab("Block") +
    ylab("Phosphate") +
    ggtitle("Phosphate by Block") +
    guides(fill=guide_legend(title="Block"))

#plot collective scatter plot separately
scatter4
#combine into a multipanel plot using borrowed multipanel function
multiplot(dhist4,dbox4,phist4,pbox4,cols=2)

#now looking at each block individually
#create a vector of colors in R
colvect<-c("aquamarine4","coral4","darkolivegreen","hotpink4","lavenderblush4","lightpink4","mediumorchid4","mediumpurple4","magenta4","violetred4","darkslategrey","burlywood4","deepskyblue4")

#set figure parameters
par(mar=c(1,1,1,1))
par(mfcol=c(4,4))

#create a scatter plot for each block
for(i in 1:13){
    subs<-subset(prob4data,block==i)
    p<-plot(x=subs$depth,y=subs$phosphate,xlab="Depth",ylab="Phosphate",col=colvect[i],
            main=paste("Block",i,sep=" "))
    assign(paste("p",i),p)
}

#recreate the combined scatter plot with altered colors for reference
plot(x=prob4data$depth,y=prob4data$phosphate,group.by=prob4data$block,col=colvect,xlab="Depth",ylab="Phosphate",main="All Blocks")

par(mfcol=c(1,1))
#create a matrix of correlation plots
pairs(prob4data,group.by=prob4data$block,col=colvect,labels=c("Block","Depth","Phosphate"))
```
\section{Problem 5}
The single most illuminating figure of this dataset is the multipanel figure of scatterplots that reveal the data of individual blocks.
\section{Appendix 1: R Code}
\subsection{Problem 4}
```{r, eval=FALSE, echo=TRUE, include=TRUE}
#get data
library(xlsx)
prob4_data1 <- read.xlsx("HW4_data.xlsx", sheetIndex = 1)
prob4_data2 <- read.xlsx("HW4_data.xlsx", sheetIndex = 2)
#combine data
prob4data<-rbind(prob4_data1,prob4_data2)

#summarize data
knitr::kable(summary(prob4data), caption="Data Summary")

library(ggplot2)

#from "http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/"
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

#begin plotting the data as a whole
#first a scatter plot
scatter4<-ggplot(prob4data,aes(x=depth,y=phosphate,colour=factor(block))) +
    geom_point() +
    xlab("Depth") +
    ylab("Phosphate") +
    ggtitle("Phosphate by Depth") +
    guides(col=guide_legend(title="Block"))

#to better understand the distributions, a histogram for both variables of interest
dhist4<-ggplot(prob4data, aes(x=depth,fill=factor(block))) +
    geom_histogram(colour="black",binwidth=5) +
    xlab("Depth") +
    ylab("Frequency") +
    ggtitle("Depth for Blocks") +
    guides(fill=guide_legend(title="Block"))
phist4<-ggplot(prob4data, aes(x=phosphate,fill=factor(block))) +
    geom_histogram(colour="black",binwidth=5) +
    xlab("Phosphate") +
    ylab("Frequency") +
    ggtitle("Phosphate for Blocks") +
    guides(fill=guide_legend(title="Block"))

#similar in purpose, a boxplot for both variables of interest
dbox4<-ggplot(prob4data, aes(x=factor(block),y=depth,group=block,fill=factor(block))) +
    geom_boxplot() +
    xlab("Block") +
    ylab("Depth") +
    ggtitle("Depth by Block") +
    guides(fill=guide_legend(title="Block"))
pbox4<-ggplot(prob4data, aes(x=factor(block),y=phosphate,group=block,fill=factor(block))) +
    geom_boxplot() +
    xlab("Block") +
    ylab("Phosphate") +
    ggtitle("Phosphate by Block") +
    guides(fill=guide_legend(title="Block"))

#plot collective scatter plot separately
scatter4
#combine into a multipanel plot using borrowed multipanel function
multiplot(dhist4,dbox4,phist4,pbox4,cols=2)

#now looking at each block individually
#create a vector of colors in R
colvect<-c("aquamarine4","coral4","darkolivegreen","hotpink4","lavenderblush4","lightpink4","mediumorchid4","mediumpurple4","magenta4","violetred4","darkslategrey","burlywood4","deepskyblue4")

#set figure parameters
par(mar=c(1,1,1,1))
par(mfcol=c(4,4))

#create a scatter plot for each block
for(i in 1:13){
    subs<-subset(prob4data,block==i)
    p<-plot(x=subs$depth,y=subs$phosphate,xlab="Depth",ylab="Phosphate",col=colvect[i],
            main=paste("Block",i,sep=" "))
    assign(paste("p",i),p)
}

#recreate the combined scatter plot with altered colors for reference
plot(x=prob4data$depth,y=prob4data$phosphate,group.by=prob4data$block,col=colvect,xlab="Depth",ylab="Phosphate",main="All Blocks")

par(mfcol=c(1,1))
#create a matrix of correlation plots
pairs(prob4data,group.by=prob4data$block,col=colvect,labels=c("Block","Depth","Phosphate"))
```













