---
title: "STAT 5014 HW5"
author: "Max McGill"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=T, cache=T, tidy.opts=list(width.cutoff=55),
                tidy=T, message=F, warning=F)
 library.warn <- library
    library <- function(package, help, pos = 2, lib.loc = NULL, character.only = FALSE,
                        logical.return = FALSE, warn.conflicts = FALSE, quietly = TRUE,
                        verbose = getOption("verbose")) {
       if (!character.only) {
          package <- as.character(substitute(package))
       }
       suppressPackageStartupMessages(library.warn(
          package, help, pos, lib.loc, character.only = TRUE,
          logical.return, warn.conflicts, quietly, verbose))}
```

\section{Problem 3}
A good figure should be visually appealing, relevant, informative, well structured, and easy to follow.
\section{Problem 4}
```{r, echo=FALSE}
#function to compute proportion of successes in a binomial vector (v) with values 1 and 0
succ<-function(v){
    sum(v)/length(v)
}

#matrix creation from assignment
set.seed(12345)
    P4b_data <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10, ncol = 10)

#apply function by column
apply(P4b_data,2,succ)
#apply function by row
apply(P4b_data,1,succ)
#we see that each column has the same proportion of success and that each row is homogeneous

#function to create vector of ten simulated coinflips from probability (p)
flips<-function(p){
  rbinom(10, 1, prob = p)
}
#vector of desired probabilities 
#wasn't sure if you still wanted only ten, since length(30:40)/100 = 11
despr<-c(.30,.31,.32,.33,.34,.35,.36,.37,.38,.39,.40)
#matrix truly desired above
coinmatr<-matrix(sapply(despr,flips),nrow=10,ncol=11)
coinmatr

#apply function by column
apply(coinmatr,2,succ)
#apply column by row
#the eleven values cause a bit of uncleanliness in these proportions
apply(coinmatr,1,succ)
```
First by column and then by row, it is seen that the provided matrix has columns with identical success proportions and rows with homogeneous compositions.
The same vector was used ten times to form the columns of the matrix.
Using the eleven proportions suggested in the previous matrix's creation, the new matrix provides individualized marginal proportions as desired, shown first by column, then by row.
\section{Problem 5}
```{r, echo=FALSE}
#get data
starch<-read.table("http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat",header=T)
#account for header, search for other sources of untidyness, begin exploration

#create basic summary for the data set as a whole
knitr::kable(summary(starch), caption="Starch Data Summary")
#repeat for individual factors
knitr::kable(summary(subset(starch,starch=="CA")), caption="Summary Table for CA")
knitr::kable(summary(subset(starch,starch=="CO")), caption="Summary Table for CO")
knitr::kable(summary(subset(starch,starch=="PO")), caption="Summary Table for PO")

#plot entire data set, treating strength as response and thickness as independent
par(mfcol=c(2,2))
plot(y=starch$strength,x=starch$thickness,ylab="Strength",xlab="Thickness",
     main="Plot of Strength by Thickness (All)")
#repeat for individual factors
plot(y=subset(starch$strength,starch$starch=="CA"),x=subset(starch$thickness,
        starch$starch=="CA"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (CA)")
plot(y=subset(starch$strength,starch$starch=="CO"),x=subset(starch$thickness,
        starch$starch=="CO"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (CO)")
plot(y=subset(starch$strength,starch$starch=="PO"),x=subset(starch$thickness,
        starch$starch=="PO"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (PO)")
#from this exploratory starting point, it can be seen that:
#   the response and independent variables appear positively linearly correlated as a whole
#   they seem less so when viewed independently of the other starches, excluding CA
#   each starch operates in different ranges of thickness and strength, which increase together
```
\section{Problem 6}
```{r, echo=FALSE}
#from assignment
#we are grabbing a SQL set from here
    # http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

    #download the files, looks like it is a .zip
    library(downloader)
    download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",dest="us_cities_states.zip")
    unzip("us_cities_states.zip", exdir="./05_R_apply_family")
    
    #read in data, looks like sql dump, blah
    library(data.table)
    states <- fread(input = "./us_cities_and_states/states.sql",skip = 19,sep = "'", sep2 = ",", header = F, select = c(2,4))
    ### YOU do the CITIES
    ### I suggest the cities_extended.sql may have everything you need
    ### can you figure out how to limit this to the 50?

#modified skip values to include all states and cities

#rename states columns
colnames(states)<-c("state","state_code")
#read in cities
    cities <- fread(input = "./us_cities_and_states/cities_extended.sql",skip = 19,sep = "'", sep2 = ",", header = F, select = c(2,4,6,8,10,12))
#rename cities columns
colnames(cities)<-c("city","state_code","zip","latitude","longitude","county")

#create vector of states
statecode<-as.vector(unique(cities$state_code))
#create vector of city counts by state
citycount<-c()
for(i in 1:52){
  citycount[i]<-length(subset(cities$state_code,cities$state_code==statecode[i]))
}

#table of number of cities by state
knitr::kable(cbind.data.frame(statecode,citycount), caption="Number of Cities by State")

#function to count occurrences of a specific letter in a sring
#from assignment
    ##pseudo code
#    letter_count <- data.frame(matrix(NA,nrow=52, ncol=26))
#    getCount <- function(letter,state_name){
#        temp <- strsplit(state_name,split=NULL)
#        count <- table(temp)
#        return(count)
#    }
#    for(i in 1:52){
#        letter_count[i,] <- xx-apply(args)
#    }

#from assignment
#https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html
#    library(ggplot2)
#    library(fiftystater)
    
#    data("fifty_states") # this line is optional due to lazy data loading
#    crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
    # map_id creates the aesthetic mapping to the state name column in your data
#    p <- ggplot(crimes, aes(map_id = state)) + 
      # map points to the fifty_states shape data
#      geom_map(aes(fill = Assault), map = fifty_states) + 
#      expand_limits(x = fifty_states$long, y = fifty_states$lat) +
#      coord_map() +
#      scale_x_continuous(breaks = NULL) + 
#      scale_y_continuous(breaks = NULL) +
#      labs(x = "", y = "") +
#      theme(legend.position = "bottom", 
#            panel.background = element_blank())
    
#    p
    #ggsave(plot = p, file = "HW5_Problem6_Plot_Settlage.pdf")
```

\section{Appendix 1: R Code}
\subsection{Problem 4}
```{r, eval=FALSE}
#function to compute proportion of successes in a binomial vector (v) with values 1 and 0
succ<-function(v){
    sum(v)/length(v)
}

#matrix creation from assignment
set.seed(12345)
    P4b_data <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10, ncol = 10)

#apply function by column
apply(P4b_data,2,succ)
#apply function by row
apply(P4b_data,1,succ)
#we see that each column has the same proportion of success and that each row is homogeneous

#function to create vector of ten simulated coinflips from probability (p)
flips<-function(p){
  rbinom(10, 1, prob = p)
}
#vector of desired probabilities 
#wasn't sure if you still wanted only ten, since length(30:40)/100 = 11
despr<-c(.30,.31,.32,.33,.34,.35,.36,.37,.38,.39,.40)
#matrix truly desired above
coinmatr<-matrix(sapply(despr,flips),nrow=10,ncol=11)
coinmatr

#apply function by column
apply(coinmatr,2,succ)
#apply column by row
#the eleven values cause a bit of uncleanliness in these proportions
apply(coinmatr,1,succ)
```
\subsection{Problem 5}
```{r, eval=FALSE}
#get data
starch<-read.table("http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat",header=T)
#account for header, search for other sources of untidyness, begin exploration

#create basic summary for the data set as a whole
knitr::kable(summary(starch), caption="Starch Data Summary")
#repeat for individual factors
knitr::kable(summary(subset(starch,starch=="CA")), caption="Summary Table for CA")
knitr::kable(summary(subset(starch,starch=="CO")), caption="Summary Table for CO")
knitr::kable(summary(subset(starch,starch=="PO")), caption="Summary Table for PO")

#plot entire data set, treating strength as response and thickness as independent
par(mfcol=c(2,2))
plot(y=starch$strength,x=starch$thickness,ylab="Strength",xlab="Thickness",
     main="Plot of Strength by Thickness (All)")
#repeat for individual factors
plot(y=subset(starch$strength,starch$starch=="CA"),x=subset(starch$thickness,
        starch$starch=="CA"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (CA)")
plot(y=subset(starch$strength,starch$starch=="CO"),x=subset(starch$thickness,
        starch$starch=="CO"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (CO)")
plot(y=subset(starch$strength,starch$starch=="PO"),x=subset(starch$thickness,
        starch$starch=="PO"),ylab="Strength",xlab="Thickness",
        main="Plot of Strength by Thickness (PO)")
#from this exploratory starting point, it can be seen that:
#   the response and independent variables appear positively linearly correlated as a whole
#   they seem less so when viewed independently of the other starches, excluding CA
#   each starch operates in different ranges of thickness and strength, which increase together
```
\subsection{Problem 6}
```{r, eval=FALSE}
#from assignment
#we are grabbing a SQL set from here
    # http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

    #download the files, looks like it is a .zip
    library(downloader)
    download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",dest="us_cities_states.zip")
    unzip("us_cities_states.zip", exdir="./05_R_apply_family")
    
    #read in data, looks like sql dump, blah
    library(data.table)
    states <- fread(input = "./us_cities_and_states/states.sql",skip = 19,sep = "'", sep2 = ",", header = F, select = c(2,4))
    ### YOU do the CITIES
    ### I suggest the cities_extended.sql may have everything you need
    ### can you figure out how to limit this to the 50?

#modified skip values to include all states and cities

#rename states columns
colnames(states)<-c("state","state_code")
#read in cities
    cities <- fread(input = "./us_cities_and_states/cities_extended.sql",skip = 19,sep = "'", sep2 = ",", header = F, select = c(2,4,6,8,10,12))
#rename cities columns
colnames(cities)<-c("city","state_code","zip","latitude","longitude","county")

#create vector of states
statecode<-as.vector(unique(cities$state_code))
#create vector of city counts by state
citycount<-c()
for(i in 1:52){
  citycount[i]<-length(subset(cities$state_code,cities$state_code==statecode[i]))
}

#table of number of cities by state
knitr::kable(cbind.data.frame(statecode,citycount), caption="Number of Cities by State")

#function to count occurrences of a specific letter in a sring
#from assignment
    ##pseudo code
#    letter_count <- data.frame(matrix(NA,nrow=52, ncol=26))
#    getCount <- function(letter,state_name){
#        temp <- strsplit(state_name,split=NULL)
#        count <- table(temp)
#        return(count)
#    }
#    for(i in 1:52){
#        letter_count[i,] <- xx-apply(args)
#    }

#from assignment
#https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html
#    library(ggplot2)
#    library(fiftystater)
    
#    data("fifty_states") # this line is optional due to lazy data loading
#    crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
    # map_id creates the aesthetic mapping to the state name column in your data
#    p <- ggplot(crimes, aes(map_id = state)) + 
      # map points to the fifty_states shape data
#      geom_map(aes(fill = Assault), map = fifty_states) + 
#      expand_limits(x = fifty_states$long, y = fifty_states$lat) +
#      coord_map() +
#      scale_x_continuous(breaks = NULL) + 
#      scale_y_continuous(breaks = NULL) +
#      labs(x = "", y = "") +
#      theme(legend.position = "bottom", 
#            panel.background = element_blank())
    
#    p
    #ggsave(plot = p, file = "HW5_Problem6_Plot_Settlage.pdf")
```