---
title: "STAT 5014 HW6"
author: "Max McGill"
date: '`r Sys.Date()`'
output: html_notebook
---

\section{Problem 2}
\subsection{Part A}
```{r, echo=FALSE, include=TRUE}
#data generation
    set.seed(12345)
    y <- seq(from=1, to=100, length.out = 1e8) + rnorm(1e8)
#a
    m<-mean(y)
#for loop
    x<-c()
for(i in 1:length(y)){
  x[i]<-(y[i]-m)^2
}
sum(x)
#timed for loop
system.time({
x<-c()
m<-mean(y)
for(i in 1:length(y)){
  x[i]<-(y[i]-m)^2
}
sum(x)
})
```
  
\subsection{Part B}
```{r, echo=FALSE, eval=TRUE, include=TRUE}
#b
#vector operations
sum((y-m)^2)
#timed vector operations
system.time({
sum((y-m)^2)
})
```
  
\section{Problem 3}
```{r, echo=FALSE, eval=TRUE, include=TRUE}
#object designation
    set.seed(1256)
    theta <- as.matrix(c(1,2),nrow=2)
    X <- cbind(1,rep(1:10,10))
    h <- X%*%theta+rnorm(100,0,0.2)
#more objects
    thetanew<-as.matrix(c(0,0),nrow=2)
    alpha<-0.05
    m<-length(h)
    tol<-0.1

#the computation
for(i in 1:m){
  while(abs(thetanew[1]-theta[1])>tol && abs(thetanew[2]-theta[2])>tol){
    thetanew[1]<-theta[1]-alpha/m*sum(thetanew[1]+thetanew[2]*X[i,2]-h[i])
    thetanew[2]<-theta[2]-alpha/m*sum((theta[1]+theta[2]*X[i,2]-h[i])*X[i,2])  
    }
}
thetanew

#for comparison
lm(h~0+X)
```
Alpha here is equal to 0.05, and the tolerance used is 0.1.
  
\section{Problem 4}
Inverting a matrix in order to solve for $\hat\beta$ would take up much more space and time than using an uninverted matrix. In this case, a potential option would be to rewrite the equation as $(X'X)\hat\beta = X'\underline{y}$ and use the software to solve for it.
  
\section{Problem 5}
```{r, echo=FALSE, eval=TRUE, include=TRUE}
#set values
  set.seed(12456) 
    
    G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
    R <- cor(G) # R: 10 * 10 correlation matrix of G
    C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
    id <- sample(1:16000,size=932,replace=F)
    q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
    A <- C[id, -id] # matrix of dimension 932 * 15068
    B <- C[-id, -id] # matrix of dimension 15068 * 15068
    p <- runif(932,0,1)
    r <- runif(15068,0,1)
    C<-NULL #save some memory space
```
\subsection{Part A}
```{r, echo=FALSE, eval=TRUE, include=TRUE}
#a
#size of A
object.size(A)
#size of B
object.size(B)
#time to calculate y
system.time({
p+(A%*%solve(B)%*%(q-r))
})
```
  
\subsection{Part B}
Firstly, it would make sense to invert B, classify that as a new matrix, and save it to call in the equation, before doing anything else. Another would be to do the same thing with q-r, naming the result elsewhere and calling it in the original subtraction's place. Multiplying the inverse of B by q-r before multiplying it by A will improve speed, as well.

\subsection{Part C}
```{r, echo=FALSE, eval=TRUE, include=TRUE}
system.time({
  G<-NULL
    Bi<-solve(B)
    st<-q-r
    A%*%(Bi%*%st)+p
})
```
  
\section{Appendix 1: R Code}
\subsection{Problem 2}
```{r, echo=TRUE, eval=FALSE, include=TRUE}
#data generation
    set.seed(12345)
    y <- seq(from=1, to=100, length.out = 1e8) + rnorm(1e8)
#a
    m<-mean(y)
#for loop
    x<-c()
for(i in 1:length(y)){
  x[i]<-(y[i]-m)^2
}
sum(x)
#timed for loop
system.time({
x<-c()
m<-mean(y)
for(i in 1:length(y)){
  x[i]<-(y[i]-m)^2
}
sum(x)
})

#b
#vector operations
sum((y-m)^2)
#timed vector operations
system.time({
sum((y-m)^2)
})
```
\subsection{Problem 3}
```{r, echo=TRUE, eval=FALSE, include=TRUE}
#object designation
    set.seed(1256)
    theta <- as.matrix(c(1,2),nrow=2)
    X <- cbind(1,rep(1:10,10))
    h <- X%*%theta+rnorm(100,0,0.2)
#more objects
    thetanew<-as.matrix(c(0,0),nrow=2)
    alpha<-0.05
    m<-length(h)
    tol<-0.1

#the computation
for(i in 1:m){
  while(abs(thetanew[1]-theta[1])>tol && abs(thetanew[2]-theta[2])>tol){
    thetanew[1]<-theta[1]-alpha/m*sum(thetanew[1]+thetanew[2]*X[i,2]-h[i])
    thetanew[2]<-theta[2]-alpha/m*sum((theta[1]+theta[2]*X[i,2]-h[i])*X[i,2])  
    }
}
thetanew

#for comparison
lm(h~0+X)
```
\subsection{Problem 5}